---
title: "DC Lead Pipe Replacement"
output: 
  flexdashboard::flex_dashboard:
    navbar:
      - { title: "Data Download", icon: "fa-download", href:  "https://www.dcwater.com/sites/default/files/cleanrivers/service-line.csv", align: right }
    orientation: rows
    vertical_layout: fill
    source_code: "https://github.com/WaterDataLab/lead"
    css: "style.css"
---

```{r setup, include=FALSE}
library(flexdashboard)
library(leaflet)
library(tidyverse)
library(sf)
library(leafgl)

# US Census poverty data
pal_pov <- c("#b3cde3", "#8c96c6", "#4d004b")
pov <- read_rds(here::here("data/dc_pov.rds")) %>% 
  st_transform(4326) %>% 
  rmapshaper::ms_simplify(keep_shapes = TRUE) %>% 
  mutate(
    col = case_when(
      percent_poverty_cat == "Most advantaged" ~pal_pov[1],
      percent_poverty_cat == "Moderately advantaged" ~pal_pov[2],
      percent_poverty_cat == "Least advantaged" ~pal_pov[3]
    ),
    lab = glue::glue("<b>{percent_poverty_cat}</b> ({percent_below_fed_poverty}% below federal poverty limit)<br><i>{NAME}</i>")
  )

# colors for valueboxes
pal_box <- rcartocolor::carto_pal(12, "Bold")

# DC lead data
url <- "https://www.dcwater.com/sites/default/files/cleanrivers/service-line.csv"

# webgl
df <- read_csv(url) %>% 
  janitor::clean_names()

# clean up
df <- df %>% 
  select(premise_address, public_service_material, 
         private_service_material, longitude, latitude) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) 

# add equity data to points
# TODO: make this file available for download 
df <- df %>% 
  st_join(select(pov, NAME:percent_poverty_cat))

# color points based on lead
lead <- c("Lead_brass", "Iron with lead", "Lead and copper", 
          "Lead", "LEAD AND COPPER", "LEAD\n", 
          "Galvanized iron with lead")

# colorblind-safe palette for points
pal_pts <- rcartocolor::carto_pal(name = "Safe")
col_noinfo = "black"
col_nolead = pal_pts[4] # green
col_privatelead = pal_pts[2] # pink
col_publiclead = pal_pts[3] # yellow
col_alllead = pal_pts[12] # grey

# add colors depending on if lead is present
# green for no lead
# orange & yellow for partial lead depending on whose side it’s on
# red for full lead
# white for no information 
df <- df %>% 
  mutate(
    no_info = ifelse(
      public_service_material == "No information" | 
        private_service_material == "No information", 
      col_noinfo, NA
    ),
    no_lead = ifelse(
      ! public_service_material %in% lead & 
        ! private_service_material %in% lead &
        public_service_material != "No information" &
        private_service_material != "No information", 
      col_nolead, NA
    ),
    private_lead = ifelse(
      private_service_material %in% lead &
        ! public_service_material %in% lead, 
      col_privatelead, NA
    ),
    public_lead = ifelse(
      public_service_material %in% lead &
        ! private_service_material %in% lead, 
      col_publiclead, NA
    ),
    all_lead = ifelse(
      public_service_material %in% lead &
        private_service_material %in% lead, 
      col_alllead, NA
    ),
    NAME = str_remove_all(NAME, ", District of Columbia, District of Columbia"),
    NAME = str_remove_all(NAME, "Census Tract "),
    popup = paste0(
      "<b>", premise_address, "</b><br>",
      "<b>Public: </b>", public_service_material, "<br>",
      "<b>Private: </b>", private_service_material, "<br>",
      # "<b>Socioeconomic group: </b>", percent_poverty_cat, "<br>",
      "<b>", percent_below_fed_poverty, "%</b> below federal poverty line")
  )

# function to format percents
f_format_percent <- function(x){
  x = x * 100
  x = round(x, 1)
  x = formatC(x, digits = 1, format = "f")
  x = paste(x, "%")
  return(x)
}
```

Map
===================================== 

Row
-------------------------------------


### No lead {.value-box}

```{r}
valueBox(f_format_percent(nrow(filter(df, !is.na(no_lead)))/nrow(df)), 
         color = col_nolead,
         caption = "Pipes with No lead")
```

### Private lead {.value-box}

```{r}
valueBox(f_format_percent(nrow(filter(df, !is.na(private_lead)))/nrow(df)), 
         color = col_privatelead,
         caption = "Pipes with Private lead")
```

### Public lead {.value-box}

```{r}
valueBox(f_format_percent(nrow(filter(df, !is.na(public_lead)))/nrow(df)), 
         color = col_publiclead,
         caption = "Pipes with Public lead")
```

### All lead {.value-box}

```{r}
valueBox(f_format_percent(nrow(filter(df, !is.na(all_lead)))/nrow(df)), 
         color = col_alllead,
         caption = "Pipes with All lead")
```

### No information {.value-box}

```{r}
valueBox(f_format_percent(nrow(filter(df, !is.na(no_info)))/nrow(df)), 
         color = col_noinfo,
         caption = "Pipes with No information")
```

Row
-------------------------------------

### **Lead pipes remaining**, last updated `r paste(Sys.time())`

```{r}
# make leaflet
leaflet() %>% 
  # 4 basemaps
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addProviderTiles(provider = providers$CartoDB.DarkMatter, group = "Dark") %>%
  addProviderTiles(provider = providers$OpenStreetMap, group = "Street") %>%
  addProviderTiles(provider = providers$Esri.WorldImagery, group = "World") %>%
  addGlPolygons(
    data = pov,
    fillColor = pov$col,
    fillOpacity = 0.8#,
    # popup = pov$lab
  ) %>% 
  # no info on private and public side
  addGlPoints(
    data = filter(df, no_info == col_noinfo),
    fillColor = "no_info",
    fillOpacity = 1,
    radius = 10,
    popup = "popup",
    group = "No Information"
  ) %>%
  # no lead on public and private side
  addGlPoints(
    data = filter(df, no_lead == col_nolead),
    fillColor = "no_lead",
    fillOpacity = 1,
    radius = 10,
    popup = "popup",
    group = "No Lead"
  ) %>%
  # private lead
  addGlPoints(
    data = filter(df, private_lead == col_privatelead),
    fillColor = "private_lead",
    fillOpacity = 1,
    radius = 10,
    popup = "popup",
    group = "Private Lead"
  ) %>%
  # public lead
  addGlPoints(
    data = filter(df, public_lead == col_publiclead),
    fillColor = "public_lead",
    fillOpacity = 1,
    radius = 10,
    popup = "popup",
    group = "Public Lead"
  ) %>%
  # all lead
  addGlPoints(
    data = filter(df, all_lead == col_alllead),
    fillColor = "all_lead",
    fillOpacity = 1,
    radius = 10,
    popup = "popup",
    group = "Private & Public Lead"
  ) %>%
  flyTo(lng = mean(st_coordinates(df)[, 1]),
        lat = mean(st_coordinates(df)[, 2]), zoom = 12) %>% 
  leaflet.extras::suspendScroll() %>% 
  addLayersControl(
    baseGroups = c("Light", "Dark", "Street", "World"),
    overlayGroups = c("No Information", "No Lead", "Private Lead", 
                      "Public Lead", "Private & Public Lead"),
    options = layersControlOptions(collapsed = FALSE),
    position = "topright"
  ) %>% 
  leaflegend::addLegendImage(
     images = "legend2.png",
     labels = "",
     title = htmltools::tags$div(
       "Legend",
       style = "font-size: 16px; text-align: center; margin-bottom: 5px;"),
     position = "topright", 
     orientation = "vertical", 
     height = 233, width = 180)
```


Replacement 
===================================== 

<center>
**VALUES MADE UP FOR DEMONSTRATION PURPOSES**
</center>
<br>


Row
-------------------------------------

### **Public** replacement rate (since 2019) {.value-box}

```{r}
# value boxes
# icons: https://ionic.io/ionicons/v2
valueBox("51%", 
         icon = "ion-merge", color = col_publiclead,
         caption = "Public replacement rate (since 2019)")
```

### **Public** pipes replaced (since 2019) {.value-box}

```{r}
valueBox("1,265 pipes", 
         icon = "ion-merge", color = col_publiclead,
         caption = "Public replacement count (since 2019)")
```


### **Private** replacement rate (since 2019) {.value-box}

```{r}
valueBox("89%", 
         icon = "ion-ios-home", color = col_privatelead,
         caption = "Private replacement rate (since 2019)")
```

### **Private** pipes replaced (since 2019) {.value-box}

```{r}
valueBox("912 pipes", 
         icon = "ion-ios-home", color = col_privatelead,
         caption = "Private replacement count (since 2019)")
```

Row
-------------------------------------

### **Lead pipes remaining**, last updated `r paste(Sys.time())`

```{r}
rates <- data.frame(
  time  = rep(c(2019, 2020, 2021), 2),
  count = c(2643, 1578, 1289, 874, 546, 95),
  type  = c(rep("Public", 3), rep("Private", 3))
) %>% 
  mutate(type = factor(type, levels = c("Public", "Private")))

p <- ggplot(rates, aes(time, count, color = type)) +
  geom_line() + 
  geom_point() + 
  scale_color_manual(values = c(col_privatelead, col_publiclead), 
                     breaks = c("Private", "Public")) +
  scale_x_continuous(breaks = 2019:2021) +
  labs(x = "", y = "Lead Pipes Remaining", color = "Type")
plotly::ggplotly(p)
```



About
===================================== 

### **Why replace lead pipes?**


<iframe width="100%" height="285" src="https://www.youtube.com/embed/6DJDGtM_d5Y" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen="" data-external="1"></iframe>


#### **History**

DC Water has historic data for the public portion of the water service line from plumbing records, service installation, and some maintenance activity that was reported to DC Water or another utility agency. Where DC Water has verified the pipe material by test pit or visual observation during a public space and/or private property service line replacement, the data source will be shown as "excavation" and is accurate as of the given inspection date. All other information is based upon historic records, but has not been confirmed. The map reflects the information DC Water has available for each active customer in the District.

### **Disclaimer**

The maps provided by the District of Columbia Water and Sewer Authority (“D.C. Water”) are based on historical data, information directly provided by customers, and in some cases, information acquired during physical inspections. DC Water does not guarantee the accuracy of these records and maps, which shall be used for the sole purpose of providing property owners and residents with DC Water’s best available data regarding their private water services, and not for any commercial, legal or other use. These records will be updated constantly as D.C. Water gathers additional information. D.C. Water requests that customers provide to it records of any service line replacements performed by property owners. D.C. Water reserves the right to alter, amend or terminate at any time the display of these maps and records.  

#### **More information**

- <a href="https://www.dcwater.com/faq-page/74" target="_blank">Service Lines - Fact Sheet</a>  
- <a href="https://www.dcwater.com/Lead" target="_blank">DC Water Lead information</a>


### **Contact**

* _General inquiries_: 202-787-2000  
* _Email_: customer.service@dcwater.com  
* _Twitter_: <a href="https://twitter.com/dcwater" target="_blank">@ dcwater</a>
